---
title: "Shine with quarto"
subtitle: "The ARCA initiative: Unlocking psychological expertise for academic and business success"
author: Ottavia M. Epifania
execute: 
  echo: false
---

This course presents a brief introduction to quarto and shiny. 

[Quarto](https://quarto.org/) is *An open-source scientific and technical publishing system*. 

[Shiny](https://shiny.posit.co/) is an R package that allows for *Easy web apps for data science without the compromises -  No web development skills required*

:::{.callout-tip icon="false"}
## Best thing 

Quarto and Shiny can work with R, Python, Jupyter and other open source softwares. 

Most importantly, shiny and quarto can be integrated, such that it is possible to develop interactive documents :)
:::

## Locations

Lectures will be at the Department of Statistical Sciences in via Cesare Battisti 241 Padova, ([https://maps.app.goo.gl/JMCthS2eZqWJEnmx8](https://maps.app.goo.gl/JMCthS2eZqWJEnmx8)), in **room SC20**.

```{r}
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng = 11.885365425918087, lat = 45.40464995544754, zoom = 13) %>%
  addMarkers(lng = 11.885365425918087, lat = 45.40464995544754, popup = "Department of Statistics")
```

Lunch will be served at the University Mensa RistorEsu Nord Piovego in Viale Giuseppe Colombo 1, Padova [https://maps.app.goo.gl/Y8ta2wHzbsdn2cMF8](https://maps.app.goo.gl/Y8ta2wHzbsdn2cMF8)

```{r}
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng =  11.890467829858775, 
          lat = 45.41117121174326, zoom = 13) %>%
  addMarkers(lng =  11.890467829858775, 
             lat = 45.41117121174326, popup = "RistorEsu Nord Piovego")

```

## Time schedule 