---
title: "Shine bright like an open source app"
subititle: "Basics"
author: "Ottavia M. Epifania"
format: html
#   revealjs: 
#     page-layout: custom
#     theme: simple
# server: shiny
runtime: shiny
---

```{r include=FALSE}
library(shiny)
```


## An astonishing result

`appExamples/basic-App.R`

::: {.panel-tabset}

<!-- ## The app -->

<!-- ```{r} -->
<!-- #| panel: center -->
<!-- selectInput(inputId = "mySelection",  -->
<!--             label = h2("Select box"), -->
<!--     choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3), -->
<!--     selected = 1) -->
<!-- verbatimTextOutput("myOutput") -->
<!-- ``` -->


<!-- ```{r} -->
<!-- #| context: server -->
<!-- output$myOutput = renderPrint({ -->
<!--     paste(paste("This is my choice"), input$mySelection) -->
<!-- }) -->
<!-- ``` -->


## The code 

```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "|1-7|8-12|13"
ui = fluidPage(selectInput(inputId = "mySelection", 
                          label = h3("Select box"),
                          choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3),
                          selected = 1),
         verbatimTextOutput(
           outputId = "myOutput"
         ))
server = function(input, output) {
           output$myOutput = renderPrint({
             paste(paste("This is my choice"), input$mySelection)
           })
}
shinyApp(ui, server)
```


:::

## App layout

```{r eval=FALSE}
#| echo: true
fluidPage()   # rows with columns within 12-unit wide grid

fluidRow(      # custom the page in two columns
  column(6, ), # of width 6 out of the 12-unit
  column(6, )  # on a single row
)

sidebarLayout(     # automatic layout with
sidebarPanel(...), # sidebar and
mainPanel(...)    # main area
)

# and many more!
```

[Complete list](https://shiny.rstudio.com/gallery/) (look for "Application layout")


## The app

```{r}
#| panel: sidebar
selectInput(inputId = "dataset", # name of the input (for the server)
                  label = "Choose a dataset:", # name of the input (for the users)
                  choices = c("rock", "pressure", "cars")) # options (for  both
                                                          # users & server)
```


```{r out.width="30%"}
#| panel: fill
 plotOutput("graph" )
verbatimTextOutput("summary")

```





```{r}
#| context: server

output$graph <- renderPlot({
    if(input$dataset == "rock"){ # call the input and its options with their label
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
    plot(data[, c(1:2)])
  })

output$summary <- renderPrint({
    if(input$dataset == "rock"){
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
    summary(data[, c(1:2)])
  })

```




## Code

`appExamples/first-app.R`

::: {.panel-tabset}

## UI

```{r}
#| echo: true
#| code-line-numbers: "|1,18|2,17|3,8|9,16|4-6|10-12|13-15"
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "dataset", # name of the input (for the server)
                  label = "Choose a dataset:", # name of the input (for the users)
                  choices = c("rock", "pressure", "cars")) # options (for  both
                                                          # users & server)
    ),
    mainPanel(
      plotOutput( # define the graphical output (we're telling R that this output
        "graph"   # container must contain a plot)
      ),
      verbatimTextOutput( # define the verbatim output (we're telling R that
        "summary"   # this output container must contain a Verbatim output)
      )
    )
  )
)
```


## Server

```{r}
#| echo: true
#| code-line-numbers: "|1,22|2-11|12-21|"
server <- function(input, output){
  output$graph <- renderPlot({
    if(input$dataset == "rock"){ # call the input and its options with their label
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
    plot(data[, c(1:2)])
  })
output$summary <- renderPrint({
    if(input$dataset == "rock"){
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
    summary(data[, c(1:2)])
  })
}
```


## Together

```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::


## Be like Paganini

::: {.panel-tabset}


## Server

```{r}
#| echo: true
#| code-line-numbers: "|2-11|12-14|15-20"
server <- function(input, output){
values <- reactiveValues() # create the shelf with reactive values
  dataInput <- reactive({ # create a reactive environment (it reacts to whatever input receives)
    if(input$dataset == "rock"){ # this part is the same as before
      data <- rock
    } else if (input$dataset == "pressure" ){
      data <- pressure
    } else if (input$dataset == "cars") {
      data <- cars
    }
  })
observe({                      # ta-da, the shelf is in action. To make it works,telling R
    values$data <- data.frame(dataInput()) #  it has to be called inside an observe function,
  })                                       # we're using a reactive object
output$graph <- renderPlot({
       plot(values$data[, c(1:2)]) # call the shelf and you're all set
    })
output$summary <- renderPrint({
    summary(values$data)
})
}

```


## UI

```{r}
#| echo: true
# the UI didn't change at all
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "dataset",
                  label = "Choose a dataset:",
                  choices = c("rock", "pressure", "cars"))

    ),
    mainPanel(
      plotOutput(
        "graph"
      ),
      verbatimTextOutput(
        "summary"
      )
    )
  )
)
```



## Together

```{r}
#| echo: true
#| eval: false
shinyApp(ui, server)
```

:::

# Slow and steady wins the race

## Take things slowly and use the action buttons

There are two ways of using the action buttons:

1. Use `eventReactive()`
2. Use the shelf we have seen before and adding the action button

(Throughout the course $\rightarrow$ [solution number 2](https://psicostat.github.io/shine-bright/slides/shelf.html#/title-slide) because I'm not sure I'm good with [solution 1](https://psicostat.github.io/shine-bright/slides/eventReactive.html#/title-slide), sorry)


